---
title: "Exploratory Data Analysis"
author: "Emma Livingston"
date: "3/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(skimr)
library(readxl)
library(here)
library(janitor)
library(viridis)
```

```{r read_data}
#Read in Target Smart Data Dictionary
ts_data_dict <- read_csv(here("/data/targetsmart_datadictionary.csv"), skip = 2) %>%
  clean_names()

#Read in de-duplicated Matched file
unique_matched <- read_csv(here("/data/Smith_Capstone_Voters_Unique.csv"))

college_campuses <- read_csv(here("/data/Smith_Capstone_College_Campus.csv"))

additional_vars <- read_csv(here("/data/Smith_Capstone_Additional_Voter_File_Vars.csv"))

college_demographics <- read_csv(here("/data/college_demographics.csv"))
```

```{r}
# codes_to_keep <- c("Thomas Jefferson Uni", "Saint Anselm College", "Communty College Phl", "UniOfNorthernIowa", "Grand View Universit", "Dartmouth College", "Drake University", "Keene State College", "Simpson College", "Antioch University", "Great Bay CommColleg", "Manchester CommColle", "Plymouth State Unive", "New England College", "NHTI","OLDCommunty College", "White Mountains ComC", "Franklin Pierce Univ", "Granite State Colleg", "University of Iowa", "Grinnell University", "OLD Grinnel", "UniofNHampshireManch", "Temple", "River Valley CommCol", "Drexel University", "Southern New Hampshi", "Uni of New Hampshire", "Uni of Pennsylvania", "Iowa StateUniversity", "Central College", "Colby-Sawyer College", "UniScienceinPhilly", "La Salle", "DesMoinesCommCollege", "Rivier University")
```


```{r}
codes_to_keep <- c(4600748,
  4600752,
  4600785,
  4598681,
  4616686,
  4600760,
  4600781,
  4600756,
  4600775,
  4600761,
  4600779,
  4600766,
  4600767,
  4600776,
  4600783,
  4600768,
  4598682,
  4600770,
  4600782,
  #4587626, pledge to vote
  4600771,
  4600754,
  4600757,
  4600758,
  4600778,
  4598678,
  4600769,
  4598684,
  4600786,
  4600777,
  4598680,
  4600755,
  4646492,
  4600780,
  4616703,
  4600750,
  4600759,
  4600784)

```



```{r}
matched_w_campuses <- unique_matched %>%
  select(-contains("activistcode")) %>%
  left_join(select(college_campuses, vanid, activistcodeid, activistcodename), by = "vanid") %>%
  filter(activistcodeid %in% codes_to_keep) 
full_data <- matched_w_campuses %>%
  left_join(select(additional_vars, van_id, vb_voterbase_race:tb_income_range_prem_cd), by = c("vanid" = "van_id"))
```


```{r}
pledge_to_vote <- unique_matched %>%
  left_join(select(additional_vars,van_id, vb_voterbase_race, tb_income_range_cd, tb_income_range_prem_cd), by = c("vanid" = "van_id")) %>%
  mutate(vb_voterbase_race = ifelse(is.na(vb_voterbase_race), "Uncoded", vb_voterbase_race))
```


# Elections 

```{r voting_mechanisms}
ggplot(unique_matched, aes(x = vb_vf_g2018)) +
  geom_bar()
```


```{r vf_vote_history_codes}
#A	 Voter cast an absentee ballot
#B	 Absentee/Mail-In, non-SOS, Applied via Voter Recognition
#E	 Voter cast an early vote ballot
#F	 Early, non-SOS, Applied via Voter Recognition
#M	 Voter cast a ballot by mail
#P	 Voter cast a ballot at the polls
#Q	 Voter cast a provisional/questionable ballot
#R	 Poll, non-SOS, Applied via Voter Recognition
#S	 Questioned/Provisional, non-SOS, Applied via Voter Recognition
#Y	 Individual voted in the election
#Z	 Voted, no vote method available, non-SOS, Applied via Voter Recognition
```

```{r}
unique_matched2 <- unique_matched %>%
  mutate(vb_vf_p2010_voted = ifelse(is.na(vb_vf_p2010), 0, 1),
  vb_vf_p2012_voted = ifelse(is.na(vb_vf_p2012), 0, 1),
  vb_vf_p2014_voted = ifelse(is.na(vb_vf_p2014), 0, 1),
  vb_vf_p2016_voted = ifelse(is.na(vb_vf_p2016), 0, 1),
  vb_vf_p2018_voted = ifelse(is.na(vb_vf_p2018), 0, 1),
  vb_vf_g2010_voted = ifelse(is.na(vb_vf_g2010), 0, 1),
  vb_vf_g2012_voted = ifelse(is.na(vb_vf_g2012), 0, 1),
  vb_vf_g2014_voted = ifelse(is.na(vb_vf_g2014), 0, 1),
  vb_vf_g2016_voted = ifelse(is.na(vb_vf_g2016), 0, 1),
  vb_vf_g2018_voted = ifelse(is.na(vb_vf_g2018), 0, 1))
  
```

```{r}
elections <- unique_matched2 %>%
  select(contains("vb_vf_g20"), contains("vb_vf_p20")) %>%
  select(contains("voted")) %>%
  pivot_longer(cols = everything(), names_to = "election", values_to = "voted") %>%
  group_by(election) %>%
  summarize(voted = sum(voted)) %>%
  arrange(desc(voted))

perc_voted_g2018 <- elections$voted[1]/nrow(unique_matched)
perc_voted_g2018
```


```{r}
ggplot(elections, aes(x = election, y = voted)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90))
```

# Demographic Breakdown

```{r}
unique_matched %>%
  select(contains("gender"))

#age variables:
#vb_voterbase_age, vb_voterbase_age_bucket

#race variable: race
#4/8 try vb_voterbase_race

#gender variable: gender


ggplot(data = unique_matched, aes(x = vb_voterbase_age_bucket, fill = gender)) +
  geom_bar(position = "dodge") +
  facet_wrap(~race)

unique_matched %>%
  group_by(vb_voterbase_age_bucket) %>%
  tally()

unique_matched %>%
  filter(vb_tsmart_state %in% c("IA", "NH", "PA", "MA")) %>%
  group_by(vb_tsmart_state, race) %>%
  tally() %>%
  mutate(percent = n / nrow(unique_matched)) %>%
  arrange(vb_tsmart_state, desc(n))
```


```{r}
nrow(filter(unique_matched, race == "Caucasian", gender == "Female", vb_voterbase_age_bucket ==  "18-29"))/nrow(unique_matched)

#this means that 35% of the people who pledged to vote were young white women
```

```{r}
ggplot(data = 
         filter(pledge_to_vote, 
                vb_voterbase_age_bucket == "18-29"), 
       aes(x = vb_voterbase_age)) +
  geom_bar(bins = 12) + 
  facet_wrap(~vb_voterbase_race)
```

```{r}
library(RColorBrewer)
ggplot(data = 
         filter(unique_matched, 
                #vb_voterbase_age_bucket == "18-29",
                race == "Caucasian"), 
       aes(x = vb_voterbase_age_bucket, fill = gender)) +
  geom_bar(bins = 12, position = "dodge") +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Pledge to Vote Individuals",
       x = "Age Bucket",
       y = "Count",
       fill = "Gender")
```

```{r}
unique_matched %>%
  select(contains("state"), contains("city")) 

#state variable: vb_tsmart_state: use this one, this is where they are living
#vb_vf_source_state is the state abbreviation indicating the state where the voter file was acquired
#city variable: vb_tsmart_city

unique_matched %>%
  group_by(vb_tsmart_state) %>%
  tally() %>%
  arrange(desc(n))

unique_matched %>%
  group_by(vb_tsmart_state) %>%
  tally() %>%
  arrange(desc(n)) %>%
  ggplot(aes(x = vb_tsmart_state, y = n)) + 
  geom_text(aes(label = vb_tsmart_state), size = 2) + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(title = "Pledge-to-Vote count by State",
       x = "State",
       y = "Count")


unique_matched %>%
  group_by(vb_vf_source_state) %>%
  tally() %>%
  arrange(desc(n)) %>%
  ggplot(aes(x = vb_vf_source_state, y = n)) + 
  geom_text(aes(label = vb_vf_source_state), size = 2) + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(title = "Pledge-to-Vote count by State",
       x = "State",
       y = "Count")
```

Targeted states: Iowa, New Hampshire, Pennsylvania, Massachusetts

## Demographic breakdown by state


```{r base_plot_function}
base_plot <- function(data, title, scales = "fixed"){ 
  data %>%
ggplot(aes(x = vb_voterbase_age_bucket, fill = gender)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("#be3a34", "#004f71", "#465a01")) +
  theme(axis.text.x = element_text(angle = 90)) + 
  facet_wrap(~vb_voterbase_race, scales = scales) +
  labs(title = title, x = "Age Group", y = "Count", fill = "Gender")
}

```

```{r demographic_state}
base_plot(data = pledge_to_vote, title = "Pledge to Vote Individuals")
#Iowa
base_plot(data = filter(pledge_to_vote, vb_tsmart_state == "IA"), title = "Iowa")

#New Hampshire
base_plot(data = filter(pledge_to_vote, vb_tsmart_state == "NH"), title = "New Hampshire")

#Pennsylvania
base_plot(data = filter(pledge_to_vote, vb_tsmart_state == "PA"), title = "Pennsylvania")

#M
base_plot(data = filter(pledge_to_vote, vb_tsmart_state == "MA"), title = "Massachusetts")
```

```{r}
pledge_to_vote %>%
  ggplot(aes(x = vb_voterbase_race, fill = vb_voterbase_race)) +
  geom_bar() +
  labs(x = "Race", y = "Count", title = "Pledge-to-Vote Individuals", fill = "Race") +
  scale_fill_manual(values = c("#be3a34", "#004f71", "#465a01", "#F2A900", "#003e51", "#981d97", "#51284f")) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_blank(),
        axis.title.x = element_text(size = 12)) 
```

Notes on next steps from 4/8 mtg:

hypothesis test between sample and population means - looking at if race was representative of the population
try to connect this back to field work research - come up with some recommendations

look for everyaction variables

a lot of the field work was done at college campuses

we could pull information about what campus they were recruited from in civis, it should be an activist code

also pull in class & education

try to look at zipcode level data


what is hte problem you're working on
visualizations
what are our conclusions - what does this mean for organizing


# Zipcode breakdown

```{r}
unique_matched %>%
  filter(vb_tsmart_state %in% c("IA", "NH", "PA", "MA")) %>%
  group_by(vb_tsmart_zip) %>%
  tally() %>%
  inner_join(unique_matched, by = "vb_tsmart_zip") %>%
  select(vb_tsmart_state, vb_tsmart_zip, n) %>%
  arrange(vb_tsmart_state, desc(n)) %>%
  unique() %>% 
  View()
```

```{r urbanicity}
unique_matched %>% 
  group_by(urbanicity) %>% 
  tally() %>% 
  View()
```

```{r}
unique_matched %>% 
  group_by(vb_vf_national_precinct_code) %>% 
  tally() %>% 
  arrange(desc(n))
```

```{r}
#time to figure out what happened with massachusetts
different_states <- pledge_to_vote %>%
  filter(vb_vf_source_state != vb_tsmart_state) 

full_data_w_college_dems <- full_data %>%
  left_join(select(college_demographics, contains("activistcode"), state, college_type), by = "activistcodeid") %>%
  rename(college_state = state.y)

not_registered_in_state <- full_data_w_college_dems %>%
  filter(vb_vf_source_state != college_state)

not_registered_in_state %>%
  group_by(college_state) %>%
  tally() %>%
  arrange(desc(n))

#new hampshire
not_registered_in_state %>%
  filter(college_state == "NH") %>%
  group_by(vb_vf_source_state) %>%
  tally() %>%
  arrange(desc(n))

college_demographics %>%
  group_by(state) %>%
  summarize(avg_out_of_state = mean(out_of_state, na.rm = TRUE))

college_demographics %>%
  group_by(state) %>%
  arrange(desc(state), desc(out_of_state)) %>%
  select(activistcodename, out_of_state) %>%
  View()
```



# Mapping 

```{r}
state_abbreviations <- tibble::tribble(
          ~STATE_NAME, ~abbreviation,
          "Alabama",  "AL",
          "Alaska", "AK",
         "Arizona", "AZ",
        "Arkansas", "AR",
      "California", "CA",
        "Colorado", "CO",
     "Connecticut", "CT",
        "Delaware", "DE",
         "Florida", "FL",
         "Georgia", "GA",
          "Hawaii", "HI",
           "Idaho", "ID",
        "Illinois", "IL",
         "Indiana", "IN",
            "Iowa", "IA",
          "Kansas", "KS",
        "Kentucky", "KY",
       "Louisiana", "LA",
           "Maine", "ME",
        "Maryland", "MD",
   "Massachusetts", "MA",
        "Michigan", "MI",
       "Minnesota", "MN",
     "Mississippi", "MS",
        "Missouri", "MO",
         "Montana", "MT",
        "Nebraska", "NE",
          "Nevada", "NV",
   "New Hampshire", "NH",
      "New Jersey", "NJ",
      "New Mexico", "NM",
        "New York", "NY",
  "North Carolina", "NC",
    "North Dakota", "ND",
            "Ohio", "OH",
        "Oklahoma", "OK",
          "Oregon", "OR",
    "Pennsylvania", "PA",
    "Rhode Island", "RI",
  "South Carolina", "SC",
    "South Dakota", "SD",
       "Tennessee", "TN",
           "Texas", "TX",
            "Utah", "UT",
         "Vermont", "VT",
        "Virginia", "VA",
      "Washington", "WA",
   "West Virginia", "WV",
       "Wisconsin", "WI",
         "Wyoming", "WY"
  )

```


```{r}
library(sf)
library(tmap)
library(leaflet)



us_shapefile <- st_read(dsn = here("data/states_21basic")) %>%
  filter(!STATE_NAME %in% c("Alaska", "Hawaii", "District of Columbia")) %>%
  left_join(state_abbreviations, by = "STATE_NAME") %>%
  left_join(tally(group_by(pledge_to_vote, vb_vf_source_state)), by = c("abbreviation" = "vb_vf_source_state")) %>%
  mutate(n = ifelse(is.na(n),0, n)) %>%
  rename(`Pledge to Vote Count` = n)

us <- tm_shape(us_shapefile) +
  tm_polygons(col = "Pledge to Vote Count",
              palette = "YlGnBu",
              breaks = c(0, 50, 100, 250, 500, 1000, Inf))
us

```

```{r}
targeted_states <- us_shapefile %>%
  mutate(highlight = ifelse(abbreviation %in% c("IA", "NH", "PA"), 1, 0))

targeted_states_map <- tm_shape(targeted_states) +
  tm_polygons(col = "highlight",
              legend.show = FALSE)
targeted_states_map
  
  
  us <- tm_shape(us_shapefile) +
  tm_polygons(col = "Pledge to Vote Count",
              palette = "YlGnBu",
              breaks = c(0, 50, 100, 250, 500, 1000, Inf))
us
```


